### 2023-05-06 paper.tex
Make corrections

TODO: write a generic paragraph

Transformers can be split into two main parts: an encoder and a decoder stack. 
The encoder reads the input text converting it into a vector representation. This is
then fed into the decoder which then generates the output utilizing this vector.
It is not necessary to have both an encoder and a decoder in the transformers
and the same tasks can be performed even without one of these present.

While encoder-decoder language models existed before the transformer, it was 
the first one to use only an attention mechanism for all computation. 
The attention mechanism is a mapping for a query and key-value pairs to some output. 
Intuitively, in transformers
it is used by the model to give different weights for different words in the input when
determining what the output should be [TODO: citation needed].

@@

The three different models studied in this paper are all similar in architecture 
being mostly faithful to the original transformer as presented in 
\cite{vaswani_attention_2017}. However, each of them have a different approach
to using the decoder and encoder stacks of the transformer. GPT uses only
the decoder, BERT only the encoder and T5 both the encoder and decoder
\cite{radford_improving_nodate,devlin_bert_2019,raffel_exploring_2020}.

\section{Training}
All three of the models this paper discusses use similar approaches for 
training. The general idea is to first train a baseline model unsupervised 
using a large amount of unstructured text and then fine tuning the model
for each specific task afterwards.

@@

In the fine-tuning stage, the pre-trained model is used as a baseline. A new
model is trained using supervised learning with a task specific data set. 
The training can fine-tune all the parameters of the model, as in BERT 
\cite{devlin_bert_2019}, or an additional layer can be added and then
trained, as in GPT \cite{radford_improving_nodate}.

@@

@@

Prefix masking can be useful for example when translating. The prefix allows the
model to attend to the entire source language sentence and causal masking can be 
used only for the target language sentence prediction. \cite{raffel_exploring_2020}

@@

\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.15]{img/attention_masking.png}
  \caption{
    Different attention masking patterns seen in transformer based language models.
    Dark squares indicate that the attention mechanism can see the input $i$ at time $j$.
    Light squares indicate that the access is blocked for a given pair $i$, $j$. 
    \cite{raffel_exploring_2020}
  }
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.15]{img/lm_masking.png}
  \caption{
    Different masking patterns used within language models. Dark lines correspond to 
    fully-visible masking. Light gray lines indicate causal masking.
    \cite{raffel_exploring_2020}
  }
  \label{masking_patterns}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.1]{img/sparse_transformer.png}
  \caption{
    Sparse Transformer. Visualizing the attention masking pattern on a 6x6 image
    for two attention heads. On the left, the original transformer. In the middle,
    a sparse transformer using striding and on the right, a fixed sparse transformer.
    \cite{child_generating_2019}
  }
  \label{sparse_transformer}
\end{figure}

@@

\subsection{Bidirectional Encoder Representations from Transformers}
Bidirectional Encoder Representations from Transformers (BERT)
is a language model utilizing masking of words during training to achieve bidirectionality
in the model. \cite{devlin_bert_2019}

@@

    \draw [-] (translateout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
    \draw [-] (colaout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
    \draw [-] (stsbout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
    \draw [-] (summaryout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
  \end{tikzpicture}
  \caption{Illustrating the types of inputs and outputs. Recreated from \cite{raffel_exploring_2020}.}
\end{figure*}

@@

\subsection{Multimodal Models}
Large language models have proven themselves very capable at a variety of language related
tasks. Their usefulness quickly diminishes the further we stray from their training set.
Using multimodal models however, we can widen the area of tasks the model
can complete and even achieve positive transfer, benefitting from the variety of tasks
it is trained for \cite{driess_palm-e_2023}.

@@

\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.1]{img/palm-e-transfer.png}
  \caption{
    PaLM-E displays positive transfer in learning. The performance is better for these three
    distinct robotics domains when trained together than when trained separately for the 
    specific tasks. \cite{driess_palm-e_2023}
  }
\end{figure}

>>>>

Natural language processing has increased in popularity. As the amount of textual data
increases, companies and researchers are looking for better ways of getting insights from
these sources. Transformers offer an answer
to the ever increasing requirements.

Transformers can be split into two main parts: an encoder and a decoder stack. 
The encoder reads the input text converting it into a vector representation. This is
then fed into the decoder which then generates the output utilizing this vector.
It is not necessary to have both an encoder and a decoder in the transformers
and the same tasks can be performed even without one of these present. 
\cite{vaswani_attention_2017}

While encoder-decoder language models existed before the transformer, it was 
the first one to use only an attention mechanism for all computation. 
The attention mechanism is a mapping for a query and key-value pairs to some output. 
Intuitively, in transformers
it is used by the model to give different weights for different words in the input when
determining what the output should be. \cite{galassi_attention_2021}

@@

The three different models studied in this paper, GPT, BERT and T5, are all similar in architecture 
being mostly faithful to the original transformer as presented in 
\cite{vaswani_attention_2017}. However, each of them have a different approach
to using the decoder and encoder stacks of the transformer. GPT uses only
the decoder, BERT only the encoder and T5 both the encoder and decoder
\cite{radford_improving_nodate,devlin_bert_2019,raffel_exploring_2020}.

\section{Training}
All three of the models this paper discusses use similar approaches for 
training. The general idea is to first train a baseline model unsupervised 
using a large amount of written text and then fine tuning the model
for each specific task afterwards.

@@

In the fine-tuning stage, the pre-trained model is adapted to a specific task 
using supervised learning.
The training can fine-tune all the parameters of the model, as in BERT 
\cite{devlin_bert_2019}, or an additional layer can be added and then
trained, as in GPT \cite{radford_improving_nodate}.

@@

Some common masking patterns include fully-visible, causal and causal with prefix.
Figure \ref{masking_patterns_raw} offers a helpful visualization for these three
patterns. 
The patterns can be mixed together and different parts of the transformer can use different
masking patterns.

@@

Prefix masking, which corresponds to the rightmost masking pattern in Figure 
\ref{masking_patterns}, can be useful for example when translating. The prefix allows the
model to attend to the entire source language sentence and causal masking can be 
used only for the target language sentence prediction. For example when translating, given an input
such as \texttt{[BEGIN]The dog is red.[PAD][SEP]Koira on punainen.[END]}, prefix masking 
can allow the model to fully attend to the input up until \texttt{[SEP]}. Only after this
point does the model have to start generating an output of its own.
 \cite{raffel_exploring_2020}

@@

\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.15]{img/attention_masking.png}
  \caption{
    Different attention masking patterns seen in transformer based language models.
    Dark squares indicate that the attention mechanism can see the input $i$ at time $j$.
    Light squares indicate that the access is blocked for a given pair $i$, $j$. 
    Figure by Raffel et al. \cite{raffel_exploring_2020}
  }
  \label{masking_patterns_raw}
\end{figure}
\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.15]{img/lm_masking.png}
  \caption{
    Different masking patterns used within language models. Dark lines correspond to 
    fully-visible masking. Light gray lines indicate causal masking.
    Figure by Raffel et al. \cite{raffel_exploring_2020}
  }
  \label{masking_patterns}
\end{figure}

\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.1]{img/sparse_transformer.png}
  \caption{
    Sparse Transformer. Visualizing the attention masking pattern on a 6x6 image
    for two attention heads. On the left, the original transformer. In the middle,
    a sparse transformer using striding and on the right, a fixed sparse transformer.
    Figure by Child et al. \cite{child_generating_2019}
  }
  \label{sparse_transformer}
\end{figure}

@@

\subsection{Bidirectional Encoder Representations from Transformers}
Bidirectional Encoder Representations from Transformers (BERT)
is a language model utilizing masking of words during training to achieve bidirectionality
in the model. Bidirectionality allows the model to simultaneously take into account
both sides of the processed word, while a generic transformer can only see one. \cite{devlin_bert_2019}

@@

    \draw [-] (translateout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
    \draw [-] (colaout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
    \draw [-] (stsbout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
    \draw [-] (summaryout.west) .. controls +(left:5mm) and +(right:5mm) .. (t5.east);
  \end{tikzpicture}
  \caption{Illustrating the types of inputs and outputs. Recreated from a Figure by Raffel et al. \cite{raffel_exploring_2020}.}
\end{figure*}

@@

\subsection{Multimodal Models}
Large language models have proven themselves very capable at a variety of language related
tasks. Unfortunately, their usefulness quickly diminishes the further we stray from their training set.
Using multimodal models however, we can widen the area of tasks the model
can complete and even achieve positive transfer, benefitting from the variety of tasks
it is trained for \cite{driess_palm-e_2023}.

@@

\begin{figure}[h]
  \centering
  \includegraphics*[scale=0.1]{img/palm-e-transfer.png}
  \caption{
    PaLM-E displays positive transfer in learning. The performance is better for these three
    distinct robotics domains when trained together than when trained separately for the 
    specific tasks. Figure by Driess et al. \cite{driess_palm-e_2023}
  }
\end{figure}
